---
title: "use_fars_function"
author: "Lina Marcela Garzon Muñoz"
date: "2019-09-13"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the packagefars package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Load Packages

```{r}
library(packagefars)
```

This document describes how to use the packagefars package and the example.

packagefars package includes five functions described below with their respective examples.



## Included data

The functions of  this package using data from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System for three years (2013, 2014, 2015)

Data files included within the package are:



```{r Setup, include = FALSE}
knitr::opts_chunk$set(# Collapse output blocks
                      collapse = TRUE, 
                      comment = "#>", 
                      fig.width = 7, 
                      fig.height = 7, 
                      fig.align = "center")
# We set the working directory where the compressed files are stored within the
# package.
x <- system.file("extdata", package = "packagefars")
knitr::opts_knit$set(root.dir = x)
library(packagefars)
```

```{r}
list.files(system.file("extdata", package = "packagefars"))
```

## Functions 1: fars_read
This function receives a .csv file and returns tbl_df of the data.This function print an error message when  .csv file does not exist

```{r}
function1 <- fars_read("accident_2013.csv.bz2")
```

## Functions 2: make_filename 

This function allows you to create accident data files per year receive the file in csv.bz2 format

```{r}
make_filename(2013)
```


### Functions 3:`fars_read_years`
This function receives a list of years (one or several years) and calls the make_filename() function for each year and then fills those files with  the corresponding year data from the main data set.


```{r}
fars_read_years(years = 2013)
```

### Function 4 `fars_summarize_years`

Receive a list of years (one or several years) . This function receives a set of data with the month and year columns, count the number number of fatalities by month for each given yea.

```{r}
# Creates a table by month & year with the total number of fatalities in 2013 y 2014.
data_2013 <- fars_summarize_years(c(2013,2014))
data_2013
```

### Function 5  `fars_map_state` 

This function creates a map with the car fatalities in a US state and a specific year.

```{r}
# Plots the car fatalities for the state of Florida (12) in 2015. 
fars_map_state(12, 2015)
```




